An Imaging Search for Post-Main-Sequence Planets of Sirius B

### 2021-11-10 paper/paper.tex
more papepr revisions and bib fixes

The reduced images do not show consistent or significant evidence for a substellar companion. The STIM probability maps for the 2020-11-21 and 2020-11-28 epochs suggest evidence for some blobs $\sim$\qty{0.3}{\au} (\ang{;;0.13}; 1.6 FWHM) from the center. The February epoch also shows a blob at a similar separation in the reduced image which does not appear in the STIM map. The lack of statistical evidence in the February epoch and the significance maps as well as the proximity to the central star both reduce the probability of these blobs being true companions. Nonetheless, we estimated astrometry for blobs from each epoch (\Cref{tbl:astrometry}) and tried fitting Keplerian orbits using the ``Orbits for the Impatient'' algorithm \citep[OFTI;][]{bluntOrbitsImpatientBayesian2017a}. We generated $10^4$ orbits, none of which constrained the points from each epoch (Appendix~\ref{sec:orbits}). This implies non-Keplerian motion and we take this as direct evidence against the blobs being substellar companions of any kind. We considered the possibility that the blobs are scattered light from a circumstellar debris disk, but this is highly unlikely given the brightness of the blob and the lack of IR excess that such a massive disk would radiate \citep{skemerSiriusImagedMidinfrared2011}. The signal can simply be explained as residual starlight not removed during PSF subtraction.

@@

Post-MS planetary evolution has historically been limited to theoretical work. Recently, though increasingly strong and diverse observational constraints, including direct detections, have invigorated the field. We set out in this work to study the nearby Sirius system for post-MS companions around Sirius B. The Sirius system is one of the most well studied in history, with Sirius B being the target of companion searches from the visible to the IR. It is highly unlikely a first-generation planet survived post-MS evolution, but no previous imaging efforts have directly addressed post-MS formation pathways in their analyses. In this work, we presented high-contrast images of Sirius B in the near-IR. Our 5$\sigma$ sensitivity limits are the best that have been reached for Sirius B so far, reaching \num{20.4} \Lp absolute magnitude at \textgreater\qty{1}{\au}. We consider multiple planetary formation pathways yielding ages between \qtyrange{126}{225}{\mega\year} and explore the effects of enriched metallicity. We translate our sensitivity limits using the ATMO2020 and Sonora Bobcat grids to constrain our mass detection limits to \qtyrange{0.7}{1.1}{\jupitermass} at \textgreater\qty{1}{\au}. Our observations also show how the high precision of the parameters of the Sirius system directly benefits the sensitivity to planets. For example, the $\sim$4\% relative uncertainty of Sirius B's age translates to mass detection limit uncertainty below \qty{0.1}{\jupitermass}. Despite the high sensitivity of this study, we found no significant evidence for a companion around Sirius B, consistent with previous results.

Although our observations yield no detections, we illustrate the capability of modern high-contrast instrumentation, even without coronagraphy, to reach strong detection limits. Our strong detection limits benefit directly from the precise stellar characterization of the Sirius system, as well as the proximity and brightness of Sirius B. With laser guide stars \citep[LGS; e.g.,][]{vandamKeckObservatoryLaser2006,baranecRoboAO2FacilityRapid2018}, the limiting magnitude for sufficient AO performances, increases significantly ($m^R\lesssim$ \num{19}). We suspect future work using LGS AO is capable of studying white dwarf systems and white dwarf binaries at the sub-AU and sub-Jupiter-mass scales \citep{holberg25ParsecLocal2016}. Such observations could significantly improve our theories of planetary formation and stellar evolution beyond the MS.

@@

\software{
ADI.jl \citep{lucasADIJlJulia2020},
astropy \citep{astropycollaborationAstropyCommunityPython2013,astropycollaborationAstropyProjectBuilding2018},
Julia \citep{bezansonJuliaFreshApproach2017},
numpy \citep{harrisArrayProgrammingNumPy2020},
scikit-image \citep{waltScikitimageImageProcessing2014},
}

@@

We found multiple interesting blobs in the reduced data that were not statistically significant. Nonetheless, we tried fitting Keplerian orbits using OFTI to determine the feasibility of the blobs being real companions. We began by estimating the astrometry of the blobs by eye in reduced data (\Cref{tbl:astrometry}, \Cref{fig:prov-orbit}). We tried simulating $10^4$ orbits via rejection sampling with OFTI but no generated orbit was able to constrain all three points. Overall we determine these blobs are not real companions and are most likely systematic noise in the stellar PSF.

\begin{deluxetable}{ccc}[h!]
    \tablecaption{Provisional astrometry for a blob of interest from each epoch. The separation and offset are in relation to Sirius B. The uncertainties are derived from the FWHM of the PSF from each epoch.
    \label{tbl:astrometry}}
    \tablehead{
        \colhead{Date observed} &
        \colhead{offset (\unit{\milliarcsecond})} &
        \colhead{PA (\unit{\degree})}
    }
    \startdata
    2020-02-04 & $123\pm 40$ & $-128\pm 20$  \\
    2020-11-21 & $119\pm 38$ & $68\pm 18$  \\
    2020-11-28 & $132\pm 41$ & $37\pm 21$  \\
    \enddata
\end{deluxetable}

\begin{figure}[h!]
    \centering
    \figurenum{13}
    \plotone{orbit_frames}
    \caption{Provisional astrometry (white circles) displayed on collapsed and derotated residuals from each epoch. Each frame was cropped to the inner $\sim$\qty{0.7}{\au} (\ang{;;0.25}) and the inner two FWHMs have been masked out. The width of the circles represents the measurement uncertainty.}
    \label{fig:prov-orbit}
\end{figure}

>>>>

The reduced images do not show consistent or significant evidence for a substellar companion. The STIM probability maps for the 2020-11-21 and 2020-11-28 epochs suggest evidence for some blobs $\sim$\qty{0.3}{\au} (\ang{;;0.13}; 1.6 FWHM) from the center. The February epoch also shows a blob at a similar separation in the reduced image which does not appear in the STIM map. The lack of statistical evidence in the February epoch and the significance maps as well as the proximity to the central star both reduce the probability of these blobs being true companions. Nonetheless, we estimated astrometry for blobs from each epoch (\Cref{tbl:astrometry}) and tried fitting Keplerian orbits using the ``Orbits for the Impatient'' algorithm \citep[OFTI;][]{bluntOrbitsImpatientBayesian2017a} using the open-source \texttt{orbitize} Python package \citep{bluntOrbitizeComprehensiveOrbitfitting2020}. We generated $10^4$ orbits, none of which constrained the points from each epoch (Appendix~\ref{sec:orbits}). This implies non-Keplerian motion and we take this as direct evidence against the blobs being substellar companions of any kind. We considered the possibility that the blobs are scattered light from a circumstellar debris disk, but this is highly unlikely given the brightness of the blob and the lack of IR excess that such a massive disk would radiate \citep{skemerSiriusImagedMidinfrared2011}. The signal can simply be explained as residual starlight not removed during PSF subtraction.

@@

Post-MS planetary evolution has historically been limited to theoretical work. Recently, though, increasingly strong and diverse observational constraints, including new detections, have invigorated the field. We set out in this work to study the nearby Sirius system for post-MS companions around Sirius B. The Sirius system is one of the most well studied in history, with Sirius B being the target of companion searches from the visible to the IR. It is highly unlikely a first-generation planet survived post-MS evolution, but no previous imaging efforts have directly addressed post-MS formation pathways in their analyses. In this work, we presented high-contrast images of Sirius B in the near-IR. Our 5$\sigma$ sensitivity limits are the best that have been reached for Sirius B so far, reaching \num{20.4} \Lp absolute magnitude at \textgreater\qty{1}{\au}. We consider multiple planetary formation pathways yielding ages between \qtyrange{126}{225}{\mega\year} and explore the effects of enriched metallicity. We translate our sensitivity limits using the ATMO2020 and Sonora Bobcat grids to constrain our mass detection limits to \qtyrange{0.7}{1.1}{\jupitermass} at \textgreater\qty{1}{\au}. Our observations also show how the high precision of the parameters of the Sirius system directly benefits the sensitivity to planets. For example, the $\sim$4\% relative uncertainty of Sirius B's age translates to mass detection limit uncertainty below \qty{0.1}{\jupitermass}. Despite the high sensitivity of this study, we found no significant evidence for a companion around Sirius B, consistent with previous results.

Although our observations yield no detections, we illustrate the capability of modern high-contrast instrumentation, even without coronagraphy, to reach strong detection limits. Our detection limits benefit directly from the precise stellar characterization of the Sirius system, as well as the proximity and brightness of Sirius B. With laser guide stars \citep[LGS; e.g.,][]{vandamKeckObservatoryLaser2006,baranecRoboAO2FacilityRapid2018}, the limiting magnitude for sufficient AO performances, increases significantly ($m^R\lesssim$ \num{19}). We suspect future work using LGS AO is capable of studying nearby white dwarf systems at the sub-AU and sub-Jupiter-mass scales \citep{holberg25ParsecLocal2016}. Such observations could significantly improve our theories of planetary formation and stellar evolution beyond the main sequence.

@@

\software{
ADI.jl \citep{lucasADIJlJulia2020},
astropy \citep{astropycollaborationAstropyCommunityPython2013,astropycollaborationAstropyProjectBuilding2018},
Julia \citep{bezansonJuliaFreshApproach2017},
numpy \citep{harrisArrayProgrammingNumPy2020},
orbitize \citep{bluntOrbitizeComprehensiveOrbitfitting2020},
scikit-image \citep{waltScikitimageImageProcessing2014},
}

@@

We found multiple interesting blobs in the reduced data that were not statistically significant. Nonetheless, we tried fitting Keplerian orbits using OFTI \citep{bluntOrbitsImpatientBayesian2017a} to determine the feasibility of the blobs being astrophysical companions. We began by estimating the astrometry of the blobs by eye in the reduced data (\Cref{tbl:astrometry}, \Cref{fig:prov-orbit}). We tried simulating $10^4$ orbits via rejection sampling with OFTI, but none of the generated orbits contained all three points in their solution. Overall we determined these blobs are not companions and are most likely systematic noise from the stellar PSF.

\begin{deluxetable}{ccc}[h!]
    \tablecaption{Provisional astrometry for a blob of interest from each epoch. The separation and offset are in relation to Sirius B. The uncertainties were derived from the FWHM of the PSF from each epoch.
    \label{tbl:astrometry}}
    \tablehead{
        \colhead{Date observed} &
        \colhead{offset (\unit{\milliarcsecond})} &
        \colhead{PA (\unit{\degree})}
    }
    \startdata
    2020-02-04 & $123\pm 40$ & $-128\pm 20$  \\
    2020-11-21 & $119\pm 38$ & $68\pm 18$  \\
    2020-11-28 & $132\pm 41$ & $37\pm 21$  \\
    \enddata
\end{deluxetable}

\begin{figure}[h!]
    \centering
    \figurenum{13}
    \plotone{orbit_frames}
    \caption{Provisional astrometry (white circles) displayed on collapsed and derotated residual frames from each epoch. Each frame was cropped to the inner $\sim$\qty{0.7}{\au} (\ang{;;0.25}) and the inner two FWHMs have been masked out. The width of the circles represents the measurement uncertainty.}
    \label{fig:prov-orbit}
\end{figure}
